---
title: "MONITORING PERSONAL ACTIVITIES"
author: "Andreia Vieira"
date: "18/08/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction


Evaluation of an individual's data collected at 5-minute intervals during the day, including the number of steps performed. Data were collected in 2012, during the months of October and November.


#### Data characteristic

Dataset: Activity Monitoring Data  (.CSV)
(https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip) [52K]
Total: 17,568 observations  


#### Variables:

steps: Number of steps taking in a 5-minute interval (missing values are coded as NA)
date: The date on which the measurement was taken in YYYY-MM-DD format
interval: Identifier for the 5-minute interval in which measurement was taken



### Data analysis


#### 1) Loading and/or preprocessing the data

```{r}

 library(ggplot2)
 library(plyr)


temp <- tempfile()
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip",temp, mode="wb")
unzip(temp, "activity.csv" )
Activity_Monitoring <- read.csv("activity.csv", sep=",", header=TRUE)
str(Activity_Monitoring)


```
###### Data clean and format

```{r}
data_clean_activity<-Activity_Monitoring[!is.na(Activity_Monitoring$steps),] #15264 obs. of  3 variables
data_clean_activity <-data_clean_activity [order(data_clean_activity$interval),] # order
data_clean_activity$date<- as.POSIXct(data_clean_activity$date, format="%Y-%m-%d") 
str(data_clean_activity)


```
#### 2) Histogram of the total number of steps taken each day

###### Sum steps

```{r}

sum_steps <- aggregate(data_clean_activity$steps ~ data_clean_activity$date, FUN=sum)
colnames(sum_steps)<- c("date", "steps")
sum_steps <-sum_steps [order(sum_steps$steps),] # order
head(sum_steps,10)

hist(sum_steps$steps, col="darkblue", border="black", xlab="Steps", main="Total Steps per Day")

```

#### 3) Average and median  - number of steps taken each Day

###### Average obtain sum steps daily: mean and median
```{r}

mean_steps <- aggregate(data_clean_activity$steps ~ data_clean_activity$date, FUN=mean)
colnames(mean_steps)<- c("date", "steps")
#mean_steps # Average each day

median_steps <- aggregate(data_clean_activity$steps ~ data_clean_activity$date, FUN=median)
colnames(median_steps)<- c("date", "steps")
#median_steps  # median each day


# Average obtain sum steps daily

mean_total_steps <- mean(sum_steps$steps)
median_total_steps <- median(sum_steps$steps)
mean_total_steps 
median_total_steps 
# mean 10766.19 and median 10765

```
###### Average steps to all interval

```{r}

mean_total_steps <- mean(data_clean_activity $steps) 
mean_total_steps


```


#### 4) Time series plot of the average number of steps taken (daily)

```{r}

mean_steps_day <- aggregate(data_clean_activity$steps ~ data_clean_activity$date, FUN=mean)
colnames(mean_steps_day)<- c("date", "steps")
head(mean_steps_day,10)

# Plot Time series  x=date

ggplot(mean_steps, aes(x = date, y = steps) , xlab="Date" , ylab="Mean Steps") + geom_hline(yintercept=mean(mean_steps_day$steps ), col="red")+ geom_line() + ggtitle("Average Number of Steps per Date")

```


#### 5) The 5-minute interval that, on average, contains the maximum number of steps

###### First Average Number of Steps per Interval. After interval that, on average, contains the maximum number of steps.
```{r}

mean_steps_interval_5min <- aggregate(data_clean_activity$steps ~ data_clean_activity$interval, FUN=mean)
colnames(mean_steps_interval_5min)<- c("interval", "steps")
head(mean_steps_interval_5min,10)


# Plot Time series  x= 5-minute interval

ggplot(mean_steps_interval_5min, aes(x = interval, y = steps) , xlab="Interval" , ylab="Mean Steps") + geom_line() + ggtitle("                  Average Number of Steps per Interval")



```

###### Maximum number of steps

```{r}

maxSteps_interval_order<-mean_steps_interval_5min[order(-mean_steps_interval_5min $steps, mean_steps_interval_5min $interval),]

#maximum

head(maxSteps_interval_order, n=1)

#All interval
head(maxSteps_interval_order,10) # all 


# The interval that contains the maximum number of steps is 835.  In this interval, the average of steps was 206.1698

```

#### 6) Code to describe and show a strategy for imputing missing data

###### The mean can distort the results due to the amount of outliers -  2304 records. So, the outliers will test by mean and mediana.  However, the substitution by the media presented a behavior similar to the clean data.

###### Calculation of median
```{r}

replace_median <- aggregate(data_clean_activity $steps ~ data_clean_activity$interval, FUN=median)
colnames(replace_median)<- c("interval", "steps")
head(replace_median,10)

```
###### Calculation of mean

```{r}

#Calculation of mean
replace_mean <- aggregate(data_clean_activity $steps ~ data_clean_activity $interval, FUN=mean)
colnames(replace_mean)<- c("interval", "steps")
head(replace_mean,10)


```
###### Filter NA

```{r}
data_for_clean <- Activity_Monitoring [is.na(Activity_Monitoring$steps),] # all outliers (read csv - line 47) 
data_for_clean$date<- as.POSIXct(data_for_clean$date, format="%Y-%m-%d") 
data_for_clean<- data_for_clean[order(data_for_clean$interval) ,]

str(data_for_clean)


```
###### Replace NA by median

```{r}

data_clean_median<-data_for_clean
m=nrow(data_clean_median) #  registros for clean
n=nrow(replace_median) # median each intervals  
for(i in 1:m)  #1 ate nrow(data_for_clean) 
       for(j in 1:n)  #1 ate nrow(replace_median)
                 if (data_clean_median$interval[i]==replace_median$interval[j]) {
                     data_clean_median$steps [i]= replace_median$steps[j]  
                     break
                }
data_replace_median<-data_clean_median
str(data_replace_median)





```
###### Replace NA by mean 

```{r}

data_clean_mean<-data_for_clean
m=nrow(data_clean_mean) #  registros for clean
n=nrow(replace_mean) # median each intervals  
for(i in 1:m)  #1 ate nrow(data_for_clean_mean) 
       for(j in 1:n)  #1 ate nrow(replace_median)
                 if (data_clean_mean$interval[i]==replace_mean$interval[j]) {
                     data_clean_mean$steps [i]= replace_mean$steps[j]  
                     break
                }
data_replace_mean<-data_clean_mean
str(data_replace_mean)





```
##### 7) Histograma com atributos  NA substuídos with media, before too test median

###### Replace by  median
```{r}

data_clean_and_NA_replace_median<-rbind(data_clean_activity, data_replace_median)
sum_steps_after_replace<-data_clean_and_NA_replace_median
sum_steps_after_replace <- aggregate(sum_steps_after_replace$steps ~ sum_steps_after_replace$date, FUN=sum)
colnames(sum_steps_after_replace)<- c("interval", "steps")
sum_steps_after_replace <-sum_steps_after_replace[order(sum_steps_after_replace$steps),] # order
head(sum_steps_after_replace,10)

# histogram initial - line 73
hist(sum_steps$steps, col="darkblue", border="black", xlab="Steps", main="Total Steps per Day")

#mean BEST RESULT
hist(sum_steps_after_replace$steps, col="red", border="black", xlab="Steps", main="Total Steps per Day") 

```

###### Replace by mean (best result)

```{r}

data_clean_and_NA_replace_mean<-rbind(data_clean_activity, data_replace_mean)
sum_steps_after_replace<-data_clean_and_NA_replace_mean
sum_steps_after_replace <- aggregate(sum_steps_after_replace$steps ~ sum_steps_after_replace$date, FUN=sum)
colnames(sum_steps_after_replace)<- c("interval", "steps")
sum_steps_after_replace <-sum_steps_after_replace[order(sum_steps_after_replace$steps),] # order
head(sum_steps_after_replace,10)

# histogram initial - line 73
hist(sum_steps$steps, col="darkblue", border="black", xlab="Steps", main="Total Steps per Day")

#mean BEST RESULT
hist(sum_steps_after_replace$steps, col="red", border="black", xlab="Steps", main="Total Steps per Day") 




```
##### 8) Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends


###### category: weekend or weekday: NA replaced media

```{r}


category_day<-data_clean_and_NA_replace_mean # data with NA replaced media

category_day$date <- as.Date(category_day$date)
category_day$days=weekdays(as.Date(category_day$date))


#replace weekend and weekday
category_day$days<- ifelse(category_day$days == "sabado" | category_day$days == "domingo", "weekend", "weekday")
head(category_day,10)

mean_category_day <- aggregate(category_day$steps ~ category_day$interval + category_day$days, FUN=mean, data=category_day)
##mean_category_day<- as.POSIXct(mean_category_day$date, format="%Y-%m-%d") 
colnames(mean_category_day)<- c("interval","days","steps")
head(mean_category_day,10)

library(lattice)
xyplot(mean_category_day$steps ~ mean_category_day$interval | mean_category_day$days, data=mean_category_day, type="l",    main="Mean Steps per Interval Based on Type of Day",   ylab="Mean Number of Steps", xlab=" Daily Interval")

```
###### category: weekend or weekday: delete dataset with NA 

```{r}

category_day<-data_clean_activity # data without NA 

category_day$date <- as.Date(category_day$date)
category_day$days=weekdays(as.Date(category_day$date))


#replace weekend and weekday
category_day$days<- ifelse(category_day$days == "sabado" | category_day$days == "domingo", "weekend", "weekday")
head(category_day,10)

mean_category_day <- aggregate(category_day$steps ~ category_day$interval + category_day$days, FUN=mean, data=category_day)
##mean_category_day<- as.POSIXct(mean_category_day$date, format="%Y-%m-%d") 
colnames(mean_category_day)<- c("interval","days","steps")
head(mean_category_day,10)

library(lattice)
xyplot(mean_category_day$steps ~ mean_category_day$interval | mean_category_day$days, data=mean_category_day, type="l",    main="Mean Steps per Interval Based on Type of Day",   ylab="Mean Number of Steps", xlab=" Daily Interval")

```

###### Conclusion item 8

According to the graphs, on weekends the number of activities is greater when compared to the other group (weekday). This was observed in the data set where NAs were replaced by the mean and in the dataset where NAs were removed.







